---
title: "Homework 2"
author: "Justin Li"
date: "2022-10-08"
output: github_document
---

```{r setup, include=FALSE}
library(nycflights13)
library(dplyr)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

What time of day should you fly if you want to avoid delays as much as possible? 
Does this choice depend on anything? Season? Weather? Airport? Airline? Find 
three patterns (“null results” are ok!). Write your results into Rmarkdown. 
Include a short introduction that summarizes the three results. Then, have a 
section for each finding. Support each finding with data summaries and 
visualizations. Include your code when necessary.
```{r}
flights %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(hour) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  arrange(desc(proportion_arrived_on_time)) 
  
```

If you want to avoid delays as much as possible, you should fly at 7 am as it 
has the highest proportion of flights that arrived on time. 

## Airlines ##
```{r}
flights %>%
  left_join(airlines, by = "carrier") %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(name) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  arrange(desc(proportion_arrived_on_time))

```

As shown in the chart above, the airlines does seem to indicate your likelihood
of arriving on time. For example, American Airlines had a 75.8% proportion of 
flights that arrived on time while still having a total of 3348 flights. On the
other hand ExpressJet Airlines had a 65.1% proportion of flights that arrived
on time while having a total of 2370 flights. This seems to indicate that 
certain airlines are better than others in terms of proportion of flights that
arrived on time. Therefore, you should choose carefully about the airline when
flying out at 7 am. 

## Origin Airport ##
```{r}
flights %>%
  left_join(airports, by = c("origin" = "faa")) %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(name) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  arrange(desc(proportion_arrived_on_time)) 
  
```

As New York only has 3 airports, and each have a very similar proportion of 
flights that arrived on time, there seems to be no relationship between the
origin airport and proportion of flights that arrived on time. Therefore, 
your choice of which airport to fly out of doesn't do anything to improve
the odds of you arriving on time at 7 am. 

## Weather (visibility) ##
```{r}
flights %>%
  left_join(weather, by = c("year", "month", "day", "hour", "origin")) %>%
  filter(hour == 7) %>%
  mutate(on_time = if_else(condition = arr_delay <= 0,
                           true = T,
                           false = F,
                           missing = F)) %>%
  group_by(visib) %>%
  summarise(proportion_arrived_on_time = mean(on_time),
            number_flights = n()) %>%
  ggplot(aes(visib, proportion_arrived_on_time)) + geom_point()
```

There seems to be a relationship between visibility and proportion of flights
that arrived on time, if we ignore the data less than visibility of 1.25 miles.
It seems like as the visibility increases, the better the proportion. Therefore,
even at the 7 am, you'll want to fly when the visibility is high. 
